Cannibalism among the supermassive black holes

### 2018-05-17 paper.tex
Made some of the changes to paper that Jerry marked up.  Also added folded for papers that we reference.

\section{Introduction}\label{sec:introduction}
A majority of all galaxies with masses $>10^8\msun$ at $z=0$ are believed to host super-massive black holes (SMBHs) in their cores. In a $\Lambda$CDM universe, galaxies at the high end of the galaxy-mass function are merger products of tens to hundreds of such massive galaxies. Even if only a small fraction of merging galaxies hosted an SMBH in their core at the time of merging, it is reasonable to assume that, during a Hubble time of continuous growth, these galaxies will have substantial phases in which they contain two or more SMBHs at the same time.

Here we look at the merger history of four exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs diffuse into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with $10^{12}<M/\msun < 10^{13}$. Galaxies with lower masses  have too few mergers with SMBH hosting galaxies. Galaxies with higher masses tend to be too extended, making dynamical friction processes inefficient and hence failing to drive SMBHs into the host galaxy core.

@@

Starting with this data, we plotted the galaxy masses as a function of time.  Since the raw data was given in terms of redshift, we converted redshift to time with the help of a 'cosmology calculator', courtesy of Ned Wright and James Schombert (\url{http://www.astro.ucla.edu/~wright/CosmoCalc.html}).  We chose four galaxies to run our simulations.  The galaxies were chosen using the following basic criteria:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($0 < z < 4$).
\item They had to have a central SMBH.
\item They had to have at least XXX orbiting black holes.
\item We ordered the galaxies by decreasing mass at $z=0$, and starting with the largest tier of mass (~$10^{14} M_{\odot}$), we spaced their masses by approximately 1-2 orders of magnitude.  We chose the four largest galaxies that also satisfied the above criteria.
\end{itemize}

@@

For our simulations, the Ostriker-Stone profile (\citet{2015ApJ...806L..28S}), which is a three-parameter potential-density pair, was used.  In order to fully parametrize this profile so as to use it in our simulations, we needed to match the stellar and halo mass data from the Kulier et al simulations, to the velocity dispersion at the galaxy's effective radius.  In Kulier's simulations, a galaxy's effective radius $R_{e}$ and velocity dispersion $\sigma(R_e)$ at the effective radius are:
\begin{equation} \label{re}
R_{e} = 2.5 kpc\left(\frac{M_*}{10^{11}M_{\odot}}\right)^{0.73}(1+z)^{-0.98}
\end{equation}
\begin{equation} \label{sig}
\sigma(R_{e}) = 190km/s\left(\frac{M_{*}}{10^{11}M_{\odot}}\right)^{0.2}(1+z)^{0.47}
\end{equation}

Given the galaxy's stellar mass $M_{*}$ at redshift z from Kulier's simulations, we can calculate the effective radius and velocity dispersion.  We can then equate these values to the analytic expressions for $\sigma(R_{e})$ in the Ostriker-Stone profile.  Using a core radius $r_c$ = 100, we can solve for the halo radius $r_h$ using a simple recursive Newton method.  Whether $\sigma_{near}$ or $\sigma_{far}$ is used is determined by whether $R_e$ is less than or greater than $\sqrt{r_c r_h}$.

>>>>

\section{Introduction}\label{sec:introduction}
A majority of all galaxies with masses $>10^8\msun$ at $z=0$ are believed to host super-massive black holes (SMBHs) in their cores. In a $\Lambda$CDM universe, galaxies at the high end of the galaxy-mass function are merger products of numerous such massive galaxies. Even if only a small fraction of merging galaxies hosted an SMBH in their core at the time of merging, it is reasonable to assume that, during a Hubble time of continuous growth, these galaxies will have substantial phases in which they contain two or more SMBHs at the same time.

Here we look at the merger history of four exemplary galaxies across the galaxy mass spectrum extracted from a cosmological simulation of hierarchical structure formation. We investigate how, after merging with incoming galaxies, SMBHs diffuse into the cores of the hosts and interact with the resident black hole. We show that gravitational interactions of multiple SMBHs are most probable in high-mass galaxies with $10^{12} \msun < M < 10^{13} \msun$. Galaxies with lower masses  have too few mergers with SMBH hosting galaxies. Galaxies with higher masses are more extended, making dynamical friction processes less efficient and hence failing to drive SMBHs into the host galaxy core.  Several papers (\citet{2018arXiv180403143I}, \citet{2014ApJ...789..156M})have shown that if all galaxy mergers lead to black hole mergers, the resulting PTA signal would be roughly twice the current limits.  One paper (\citet{2018MNRAS.473.3410R}) has indicated that dynamical interactions among multiple black holes, which eject a non-negligible fraction of the mass, may solve this problem.  The present paper also addresses this purported dynamical solution.

@@

Starting with this data, we plotted the galaxy masses as a function of time.  Since the raw data was given in terms of redshift, we converted redshift to time.  We chose four galaxies to run our simulations.  The galaxies were chosen using the following basic criteria:
\begin{itemize}
\item The galaxies had to exist through the entire simulation ($0 < z < 4$).
\item They had to have a central SMBH.
\item They had to have at least XXX orbiting black holes.
\item We ordered the galaxies by decreasing mass at $z=0$, and starting with the largest tier of mass (~$10^{14} M_{\odot}$), we spaced their masses by approximately 1-2 orders of magnitude.  We chose the four largest galaxies that also satisfied the above criteria.
\end{itemize}

@@

For our simulations, the Stone-Ostriker profile (\citet{2015ApJ...806L..28S}), which is a three-parameter potential-density pair, was used.  It is essentially an analytic form of a finite, cored isothermal mass distribution.  In order to fully parametrize this profile so as to use it in our simulations, we needed to match the stellar and halo mass data from the Kulier et al simulations, to the velocity dispersion at the galaxy's effective radius.  In Kulier's simulations, a galaxy's effective radius $R_{e}$ and velocity dispersion $\sigma(R_e)$ at the effective radius are:
\begin{equation} \label{re}
R_{e} = 2.5 kpc\left(\frac{M_*}{10^{11}M_{\odot}}\right)^{0.73}(1+z)^{-0.98}
\end{equation}
\begin{equation} \label{sig}
\sigma(R_{e}) = 190km/s\left(\frac{M_{*}}{10^{11}M_{\odot}}\right)^{0.2}(1+z)^{0.47}
\end{equation}

Given the galaxy's stellar mass $M_{*}$ at redshift z from Kulier's simulations, we can calculate the effective radius and velocity dispersion.  We can then equate these values to the analytic expressions for $\sigma(R_{e})$ in the Stone-Ostriker profile.  Using a core radius $r_c$ = 100, we can solve for the halo radius $r_h$ using a simple recursive Newton method.  Whether $\sigma_{near}$ or $\sigma_{far}$ is used is determined by whether $R_e$ is less than or greater than $\sqrt{r_c r_h}$.
